---
title: "Week12 Assign"
author: "Nathan Lim"
date: "April 20, 2015"
output: html_document
---
The task is,

(1) Write the contents of a json file of your choosing into a mongodb database, then</br>
(2) Use R to display some of the information from the mongodb database.

* * *
```{r}
library(rjson)
library(rmongodb)
library(knitr)
```

* * *
Before I did R coding, I made a database in MongoDB by importing "new york jazz perfomance" data using "mongoimport".

Among the collections below, the data I want is "nyjazz" that has information about jazz musicians, addresses, etc.

```{r}
mongo <- mongo.create()
mongo
mongo.is.connected(mongo)
if(mongo.is.connected(mongo) == TRUE) {
        mongo.get.databases(mongo)
}

if(mongo.is.connected(mongo) == TRUE) {
        db <- "mydb"
        mongo.get.database.collections(mongo, db)
}
```

* * *

I changed the bson to list, and then changed it to a data frame.

```{r}
coll <- "mydb.nyjazz"
tmp <-  mongo.find.one(mongo,coll)
class(tmp)
mgdata <- mongo.bson.to.list(tmp)
df <- data.frame(musician=mgdata[[2]], place=mgdata[[3]], address=mgdata[[4]], time=mgdata[[5]], stringsAsFactors = FALSE)
str(df)
```

* * *
Searching jazz performances in Lincon center

```{r}
kable(df[grep('Center', df$place),])
```
