---
title: "IS607_Assign11"
author: "Nathan Lim"
date: "April 13, 2015"
output: html_document
---

The New York Times web site provides diverse APIs including article search, books, the community, etc.
Among them, I got "THE EVENT LISTINGS API" key from the web site.
What I want to do is to retrieve Jazz performances in Manhattan between 2015-04-13 and 2015-04-20

```{r, warning=FALSE, message=FALSE}
library(RCurl)
library(rjson)
library(dplyr)
library(plyr)
library(knitr)
```

Get URL, including API key and filters(Manhattan, dates between 2015-04-13 and 2015-04-20)
```{r, warning=FALSE, message=FALSE}
theURL <- "http://api.nytimes.com/svc/events/v2/listings.json?filters=category:Jazz,borough:Manhattan&date_range=2015-04-13:2015-04-20&api-key=95a41335ddef20739e14c93a36b67c0d:5:71828133"
url_json <- getURL(theURL)
data = fromJSON(url_json)
```

Select the "result" list, which is 4th from the list, and then change the result list to a data frame
```{r, warning=FALSE, message=FALSE, results="hide"}
str(data)
rdf <- data[[4]]
rdf2 <- ldply(rdf, function(x){data.frame(x)})
```

Select columns we want to know using "dplyr"
```{r}
rdf_new <-rdf2%>%
        select(event_name, venue_name, street_address, date_time_description, category)
kable(distinct(rdf_new))
